---
title: eDNA detections quality control
output: html_document
editor_options: 
  chunk_output_type: console
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "docs") })
---

<style type="text/css">
.main-container {
  max-width: 1800px !important;
  margin: auto;
}
p {
  margin: 30px 0px;
}
</style>

The table below shows the likelihood score of eDNA detections based on WoRMS distributions, species thermal envelopes built from OBIS and GBIF data, and distance to known observations in OBIS and GBIF. Identifications are done using RDP classifier, but conflicting identifications from VSEARCH are shown as well. The original data are available from <https://github.com/iobis/edna-results/tree/data/data>.

```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(dplyr)
library(purrr)
library(speedy)
library(stringr)
library(ggplot2)
library(sf)
library(glue)
library(formattable)
library(tidyr)

fish_classes <- c("Actinopteri", "Cladistii", "Coelacanthi", "Elasmobranchii", "Holocephali", "Myxini", "Petromyzonti", "Teleostei")
turtle_orders <- c("Testudines")
mammal_classes <- c("Mammalia")
st_qc <- storr::storr_rds("storr_qc")
max_distance <- 5000
coord_round <- 3
options(knitr.kable.NA = "")
```

```{r, message=FALSE, warning=FALSE, echo=FALSE}
# read eDNA results

dna_files <- list.files("output", "*DNADerivedData*", full.names = TRUE)
occurrence_files <- list.files("output", "*Occurrence*", full.names = TRUE)

dna <- map(dna_files, read.table, sep = "\t", quote = "", header = TRUE) %>%
  bind_rows() %>%
  mutate_if(is.character, na_if, "")

occurrence <- map(occurrence_files, read.table, sep = "\t", quote = "", header = TRUE) %>%
  bind_rows() %>%
  mutate_if(is.character, na_if, "") %>%
  mutate(
    aphiaid = str_replace(scientificNameID, "urn:lsid:marinespecies.org:taxname:", ""),
    species = ifelse(taxonRank == "species", scientificName, NA)
  ) %>%
  left_join(dna, by = "occurrenceID")

# extract vsearch result

occurrence <- occurrence %>%
  mutate(vsearch = str_replace(str_replace(str_replace(str_match(identificationRemarks, "VSEARCH.*;(.*).")[,2], "_", " "), "\\(", "\\\\("), "\\)", "\\\\)")) %>%
  mutate(vsearch = ifelse(str_detect(paste0(phylum, class, order, family, genus, scientificName), vsearch), NA, vsearch))

# add groups

occurrence <- occurrence %>%
  mutate(
    group = case_when(
      class %in% fish_classes ~ "fish",
      order %in% turtle_orders ~ "turtle",
      class %in% mammal_classes ~ "mammal"
    )
  )

# add qc

detections <- occurrence %>%
  filter(!is.na(species) & !is.na(decimalLongitude)) %>%
  group_by(higherGeography, species, class, family, group, aphiaid) %>%
  summarize(
    batch = paste0("b", str_replace_all(paste0(unique(sort(eventRemarks)), collapse = ","), "sequencing batch ", "")),
    occurrenceID = paste0(head(occurrenceID, 2), collapse = ", "),
    reads = sum(organismQuantity, na.rm = TRUE),
    decimalLongitude = first(decimalLongitude),
    decimalLatitude = first(decimalLatitude),
    vsearch = paste0(head(unique(na.omit(vsearch)), 2), collapse = ", ")
  ) %>%
  ungroup() %>%
  mutate(key = glue("{aphiaid}_{round(decimalLongitude, 3)}_{round(decimalLatitude, 3)}"))

unique_keys <- unique(detections$key)
qc_list <- st_qc$mget(unique_keys)
names(qc_list) <- unique_keys
qc <- bind_rows(qc_list, .id = "key")

detections <- detections %>%
  left_join(qc, by = "key")

detections <- detections %>%
  rowwise() %>%
  mutate(distance = round(distance / 1000)) %>% 
  mutate(across(where(is.character), ~replace_na(.,""))) %>%
  mutate(
    score = ifelse(
      worms, 1, ifelse(
        thermal,
        max(1 * (max_distance - distance) / max_distance, 0),
        max(0.5 * (max_distance - distance) / max_distance, 0)
      )
    )
  ) %>%
  mutate(score = round(score, 2)) %>%
  select(site = higherGeography, batch, class, family, species, vsearch, group, distance, worms, thermal, score, aphiaid, reads, occurrenceID) %>%
  arrange(site, species)

# rm("occurrence", "dna", "qc_list", "qc")
```

```{r, message=FALSE, warning=FALSE, echo=FALSE}
icon_formatter <- function() {
  formatter("span", 
    style = x ~ style(color = ifelse(x, "#a7c400", "#f57842")), x ~ icontext(ifelse(x, "ok", "remove"), "")
  )	 	 
}

distant_formatter <- function() {
  formatter("span", 
    style = ~ style(color = ifelse(distance > max_distance, "#f57842", "black"))
  )	 	 
} 

vsearch_formatter <- function() {
  formatter("span", 
    style = ~ style(color = "#f57842")
  )	 	 
} 

ft <- detections %>%
  mutate(species = glue("<a target=\"_blank\" href=\"https://www.marinespecies.org/aphia.php?p=taxdetails&id={aphiaid}\">{species}</a>")) %>%
  mutate(vsearch = ifelse(is.na(vsearch), vsearch, glue("<a target=\"_blank\" href=\" https://www.marinespecies.org/aphia.php?p=taxlist&tName={vsearch}\">{vsearch}</a>"))) %>%
  select(-aphiaid) %>%
  formattable(
    align = rep("l", 6),
    list(
      score = color_tile("#f57842", "#ffffff"),
      distance = distant_formatter(),
      worms = icon_formatter(),
      thermal = icon_formatter()
    )
  )
ft %>%
  as.datatable(escape = FALSE, rownames = FALSE, filter = "none", options = list(pageLength = 500))
```
